# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data

1. Load the CSV file using Date class.

  ```{r dataload}
  activity = read.csv("activity.csv", colClasses=c("numeric", "Date", "numeric"))
  ```

## What is mean total number of steps taken per day?

1. Make a histogram of the total number of steps taken each day. 
Aggregate the steps taken by date and plot histogram.
  ```{r histogram}
  library(lattice)
  totalsteps = aggregate(steps ~ date, data = activity, FUN = sum)
  xyplot(steps ~ date, data = totalsteps, type = "h", lwd = 5, main = "Total steps per day")
  ```

2. Calculate and report the mean and median total number of steps taken per day.
  ```{r mean}
  meansteps = mean(totalsteps$steps)
  mediansteps = median(totalsteps$steps)
  ```

  The mean and median total number of steps taken per day is **`r as.integer(meansteps)`** and **`r as.integer(mediansteps)`** respectively.

## What is the average daily activity pattern?

1. Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis).
  ```{r timeseriesplot}
  library(lattice)
  library(stats)
  avgsteps = aggregate(steps ~ interval, data = activity, FUN = mean)
  xyplot(steps ~ interval, data = avgsteps, type = "l", main = "Average Steps per 5-minute Interval")
  ```

2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

  ```{r max}
  sortedavg = avgsteps[order(avgsteps$steps),]
  maxinterval = sortedavg[length(sortedavg[[2]]), 1]
  maxsteps = sortedavg[length(sortedavg[[2]]), 2]
  ```
  
  The **`r maxinterval`**th 5-minute interval contains the maximum number, **`r as.integer(maxsteps)`**, of steps.

## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
